Using simulation to optimize the sampling design of fisheries-independent trawl surveys

### 2018-09-14 analysis/paper/sim_survey_paper.Rmd
Add abundance-at-length calculations

$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
Here, numbers-at-age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e. recruits) vary around a baseline value, $log(N_{1,y}) = log(\mu_r) + \epsilon_{y}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(\mu_Z) + \delta_{a,y}$. The error around the recruitment process was set to follow a random walk, $\epsilon_{y} \sim N(\epsilon_{y - 1}, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$. The covariance across ages and years is controlled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, age}$ and $\varphi_{\delta, year}$, respectively). This structure allows for autocorrelation in process errors across ages and years (i.e. total mortality can be made to be more similar for fish that are closer together in age and time). Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.

@@

Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean. This helps simulate depth preferences of particular fish. Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years. Spatial correlations was controlled by a smoothing ($\lambda$) and scaling parameter ($\kappa$) [here we approximate $\kappa$ from a range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years was controlled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively. The overall variance of the process was controlled by $\sigma_{\xi}^2$. In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering. Note that $\eta_{a,y,s}$ is forced to sum to one in each year to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model. Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid. A more detailed description of the space-age-year covariance is included in Appendix 1.

@@

$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{trawl}}{A_{cell}} q_a )$$
Where $n_{a,y,s}$ is the number of fish sampled by a set, $A_{trawl}$ indicates the area covered by the trawl, $A_{cell}$ is the area of a grid cell (i.e. $R^2$), and $q_a$ is the catchability coefficient of each age. Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$. The lengths of the fish sampled by the set are then simulated using the original von Bertalanffy growth curve [@Cailliet2006]:

$$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$
Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$. Depending on the number of fish caught, sub-sampling is then conducted. Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division.

@@

Ideally the parameter settings used here would be based on model fits to real data, however, conditioning the operating model described above was outside the scope of this project. Moreover, many of the parameters described above may not be identifiable or it may be difficult to obtain un-biased estimates [e.g. bias introduced by length-stratified sampling complicate the fitting of growth models; @Morgan1997]. As such, parameters were informally selected to roughly correspond to observations typically obtained by the Fisheries and Oceans Canada survey of cod in NAFO division 3Ps, Newfoundland. Though this approach is far from exact, the goal was to approximate the life-history and distribution of an actual population to simulate an arena for testing the consequences of surveys with different sampling protocol. Parameter settings and descriptions used in this case study are presented in Table \@ref(tab:par). A series of `r nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests). Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain an estimate of abundance at age for each year for each simulation, $\hat{I}_{a,y,k}$. Intra-haul correlation (i.e. intraclass correlation) of the set catch was calculated by fitting a mixed-model to the lengths data with set as a random effect [fit using the lme4 package; @Bates2015]; variance estimates from this model were then plugged into the ICC(1) formula presented in McGraw and Wong [-@Mcgraw1996]. We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance estimates from each survey:

$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})^2}{N_a N_y N_k}}$$
Where $N_a$, $N_y$, and $N_k$ are the number of years, ages and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e. catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$). All simulations and analyses were conducted using `r substr(version$version.string, 1, 15)` [@R2017] and the code has been developed into a package called `SimSurvey` to facilitate future work and collaboration (https://github.com/PaulRegular/SimSurvey).

@@

- Feedback and advice: Alejandro Buren, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Dwayne Pittman, Don Power, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
- Dave Cote for a review of a previous version of this manuscript

@@

@@

|                                   Mean asymptotic length |               $l_{\infty}$ |            120 |
|                                          Length at birth |                      $l_0$ |              5 |
|                                             Length group |                $l_{group}$ |              1 |  

>>>>

$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
Here, numbers-at-age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e. recruits) vary around a baseline value, $log(N_{1,y}) = log(\mu_r) + \epsilon_{y}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(\mu_Z) + \delta_{a,y}$. The error around the recruitment process was set to follow a random walk, $\epsilon_{y} \sim N(\epsilon_{y - 1}, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$. The covariance across ages and years is controlled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, age}$ and $\varphi_{\delta, year}$, respectively). This structure allows for autocorrelation in process errors across ages and years (i.e. total mortality can be made to be more similar for fish that are closer together in age and time). Abundance-at-age is then converted to abundance-at-length using the original von Bertalanffy growth curve [@Cailliet2006]:

$$log(l) = log(l_{\infty} - (l_{\infty} - l_0) e^{-Ka}) + \varepsilon$$

Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$. Abundance in discrete length groups, $l_{group}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance-at-age to abundance-at-length. Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.

@@

Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean. This helps simulate depth preferences of particular fish. Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years. Spatial correlations was controlled by a smoothing ($\lambda$) and scaling parameter ($\kappa$) [here we approximate $\kappa$ from a range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years was controlled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively. The overall variance of the process was controlled by $\sigma_{\xi}^2$. In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering. Note that $\eta_{a,y,s}$ is forced to sum to one in each year to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model. Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid. A more detailed description of the space-age-year covariance is included in Appendix 1. Note that the the distributions of specific length groups were not simulated because of the greater computational cost of storing spatially dis-aggregated abundance across all length groups.

@@

$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{trawl}}{A_{cell}} q_a )$$
Where $n_{a,y,s}$ is the number of fish sampled by a set, $A_{trawl}$ indicates the area covered by the trawl, $A_{cell}$ is the area of a grid cell (i.e. $R^2$), and $q_a$ is the catchability coefficient of each age. Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$. The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the [Simulate abundance] section. Depending on the number of fish caught, sub-sampling is then conducted. Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division.

@@

Ideally the parameter settings used here would be based on model fits to real data, however, conditioning the operating model described above was outside the scope of this project. Moreover, many of the parameters described above may not be identifiable or it may be difficult to obtain un-biased estimates [e.g. bias introduced by length-stratified sampling complicate the fitting of growth models; @Morgan1997]. As such, parameters were informally selected to roughly correspond to observations typically obtained by the Fisheries and Oceans Canada survey of cod in NAFO division 3Ps, Newfoundland. Though this approach is far from exact, the goal was to approximate the life-history and distribution of an actual population to simulate an arena for testing the consequences of surveys with different sampling protocol. Parameter settings and descriptions used in this case study are presented in Table \@ref(tab:par). A series of `r nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests). Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain annual estimates of total abundance ($\hat{I}_{y,k}$), abundance-at-length ($\hat{I}_{l,y,k}$) and abundance-at-age ($\hat{I}_{a,y,k}$). We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance-at-age estimates from each survey:

$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})^2}{N_a N_y N_k}}$$
Where $N_a$, $N_y$, and $N_k$ are the number of years, ages and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e. catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$). Intra-haul correlation (i.e. intraclass correlation) of set catch was also calculated by fitting a mixed-model to the lengths data with set as a random effect [fit using the lme4 package; @Bates2015]; variance estimates from this model were then plugged into the ICC(1) formula presented in McGraw and Wong [-@Mcgraw1996]. All simulations and analyses were conducted using `r substr(version$version.string, 1, 15)` [@R2017] and the code has been developed into a package called `SimSurvey` to facilitate future work and collaboration (https://github.com/PaulRegular/SimSurvey).

@@

- Feedback and advice: Alejandro Buren, Noel Cadigan, Karen Dwyer, Geoff Evans, Brian Healey, Paul Higdon, Danny Ings, Mariano Koen-Alonso, Joanne Morgan, Derek Osborne, Dwayne Pittman, Don Power, Martha Robertson, Mark Simpson, Brad Squires, Don Stansbury, Peter Upward
- Dave Cote for a review of a previous version of this manuscript

@@

|                                   Mean asymptotic length |               $l_{\infty}$ |         120 cm |
|                                          Length at birth |                      $l_0$ |           5 cm |
|                     Length group for abundance at length |            $l_{group}^{N}$ |           3 cm |

@@

|                            Length group for age sampling |                $l_{group}$ |           1 cm |  
